Fusion Moves for Correlation Clustering

### 2014-11-11 paper/egpaper_for_review.tex
made figures a bit smaller

Fusion moves as defined in \cite{Lempitsky-2010} work
in the node domain and do not work properly for 
objective functions as Eq.~\ref{eq:nodeproblem} since
the node coloring is ambiguous and has no semantic meaning (see Fig.~\ref{fig:fusion_abiguty}).
In the following, we propose a more suitable fusion move for correlation
clustering which works on the edge domain.
%
Given two proposal solutions $y'$ and $y''$,
$\mathcal{C}$ is the set of edges
which are uncut in $y'$ and $y''$.
%
\begin{align}
\breve{y}    & =  y' \cup y'' \\
\mathcal{C}  & =  \{ (i, j) \quad | \quad \breve{y}_{ij} = 0 \}
\end{align}
%
The fusion move for correlation clustering is solving Eq.~\ref{eq:edgeproblem}
with additional \emph{must-link constraints} for all edges in $\mathcal{C}$.
%
\begin{align}
  y^* &=& \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:fusion_move_a}.\\ 
  s.t & &  y_{ij} = 0 \quad \forall (i, j) \in \mathcal{C} \nonumber
\end{align}
%
By construction, solving Eq.~\ref{eq:fusion_move_a} cannot increase the energy:
\begin{align}
  \sum  w_{ij} \cdot y^*_{ij}  \leq \min \left( \sum  w_{ij} \cdot y'_{ij} ,  \sum   w_{ij} \cdot y''_{ij} \right)
\end{align}
%
Eq.~\ref{eq:fusion_move_a} can be solved efficiently, since 
we can work on a coarser graph, where all nodes which are connected
via must-link constraints are merged into single nodes.
This coarser graph is constructed by contracting all must-link edges
(see Fig.~\ref{fig:move_graph}).

@@

\subsection{Fast Optimization} \label{sec:fast_optimization}
TODO:
Here we need to proof that it is equivalent 
to optimize the contracted graph !

@@

% Global optimal solvers for multicut do not scale beyond ??? \cite{???}.
% Good approximate solvers for planar graphs exist \cite{beier_2014_cvpr,yarkony_2012_eccv} 
% but have difficulties to find good solutions for non planar graphs \cite{beier_2014_cvpr}.

% To make multicuts scalable, it is desirable to replace the graph $G$
% with a graph $\hat{G}$, which is smaller in the number of nodes and edges. 
% If $\hat{G}$ is small enough, it is feasible to use global optimal solvers.

% A trivial way to reduce the size of the graph  is edge contraction.
% Until a desirable size is reached, one might contract the edge with
% the highest weight. A cut on the contracted graph  $\hat{G}$ is always
% a valid cut on $\hat{G}$.

% On the other hand, if an edge $e$ is contracted, it will never be cut.
% If $e$ is part of the cut in a global optimal solution, we cannot find
% this cut.

% To overcome this problem  we can randomize the edge contraction,
% and repeat this multiple times and remember the best solution.

% But this would throw away all the ``not the best'' solutions (rewrite me).
% Within this work we propose the following nifty trick.
% Instead of throwing away solutions, we intersect them with the 
% best solution, and solve the multicut on the resulting
% graph.
% Therefore we fuse multiple solutions,

%\begin{algorithm}
%\begin{scriptsize}
%\caption{Fusion Based Algorithms}\label{alg:fusion}
%
%
%
%
%\begin{algorithmic}[1]
%\Procedure{Intersection-Based-Inference}{GEN,$J$,$X$}
%\State $x^0 \gets \textrm{initial state form } X$
%\State $n \gets 0 $                                 \Comment{Number of moves}
%\State $m \gets 0 $                                 \Comment{Number of moves without progress}
%\While{$m < m_{\max}$ and $n < n_{\max}$ }
%\State $n \gets n+1$ 
% \State $x'\gets GEN(x^{n-1},J,X)$                    \Comment{Generate proposal}
% \If{$J(x^{n-1}) \leq J(x')$}                       
%   \State $x^{n} \gets Fuse(x^{n-1},x',J)$             
% \Else
%   \State $x^{n} \gets Fuse(x',x^{n-1},J)$
% \EndIf
% \If{$J(x^{n}) \leq J(x^{n-1})$}
%   \State $m \gets 0$                                 \Comment{Reset counter}
% \Else
%   \State $m \gets m+1$                               \Comment{Increment counter}
% \EndIf
%\EndWhile
%\State \textbf{return} $x^n$
%\EndProcedure
%\end{algorithmic}
%\end{scriptsize}
%\end{algorithm}

%\begin{algorithm}
%\begin{scriptsize}
%\caption{Fusion Moves}\label{alg:fusion_moves}
%\scriptsize
%\begin{algorithmic}[1]
%\Require $J(x) \leq J(x')$
%\Ensure $J(\hat{x}) \leq J(x)$
%\vspace{0.2cm}
%\Procedure{Fuse}{$x,x',J$}
%\State $\bar{x} \gets \textbf{\scriptsize{Intersect}}(x,x')$  
%    \Comment{
%        \begin{minipage}{0.4\linewidth} 
%            \begin{tiny}
%            \begin{singlespace}
%                intersect
%                uncut edges in $x$ and $x'$,
%                return  
%                connected component labeling
%            \end{singlespace}
%            \end{tiny}
%        \end{minipage}
%    }
%\vspace{0.1cm}
%\State $ \tilde{G}=(\tilde{E},\tilde{V}),\tilde{W} 
%    \gets \textbf{\scriptsize{Contract}}(G,W,\bar{x})$  
%    \Comment{ 
%        \begin{minipage}{0.2\linewidth} 
%            \begin{tiny}
%            \begin{singlespace}
%                Contract all edges 
%                which are uncut in $\bar{x}$ 
%            \end{singlespace}
%            \end{tiny}
%        \end{minipage}
%    }
%\vspace{0.1cm}
%\State $\tilde{x}^*= \argmin\limits_{\tilde{x}} \sum\limits_{e_{ij} \in \tilde{E}} \tilde{w}_{ij}[\tilde{x}_i \neq \tilde{x}_j]$
%    \Comment{ 
%        \begin{minipage}{0.3\linewidth} 
%            \begin{tiny}
%            \begin{singlespace}
%                Solve the multicut objective
%                on smaller graph $\tilde{G}=(\tilde{V},\tilde{E})$
%            \end{singlespace}
%            \end{tiny}
%        \end{minipage}
%    }
%
%\State $\hat{x} \gets \textbf{\scriptsize{ProjectBack}}(\tilde{x}^*)$
%    \Comment{ 
%        \begin{minipage}[\textheight]{0.4\linewidth} 
%            \begin{tiny}
%            \begin{singlespace}
%                Translate the connected component
%                labeling $\tilde{x}^*$ of $\tilde{G}$ 
%                to a connected component labeling of $G$
%            \end{singlespace}
%            \end{tiny}
%        \end{minipage}
%    }
%\State \textbf{return} $\hat{x}$
%\EndProcedure
%\vspace{0.3cm}
%\Procedure{FUSE$_{\textrm{BASE}}$}{$x,x',J$}
%\State \textbf{return} ${\arg\min}_{\bar{x} \in \{x,x'\}}J(\bar{x})$
%\EndProcedure
%\end{algorithmic}
%\end{scriptsize}
%\end{algorithm}

%-------------------------------------------------------------------------
\subsection{Fusion Move Solver}

% \begin{center}
%     \begin{eqnarray}
%         y 
%         &=& 
%         \rho_{\hat{y}}^{-1}(
%         \argmin_{y \in MC(G_{\hat{y}})\cap \{0,1\}^{|E_{\hat{y}}|}} 
%         \sum_{ (i,j) \in E_{\hat{y}} } 
%         w^{\hat{y}}_{ij} \cdot \bar{y}_{ij} 
%     \end{eqnarray}
% \end{center}

% \begin{center}
%     \begin{eqnarray}
%       y \cdot w \leq \min(y'\cdot w,  y'' \cdot w) 
%     \end{eqnarray}
% \end{center}
Once we know how to fuse segmentations, the actual algorithm 
for correlation clustering 
is simple and similar to the fusion move algorithm of Lempitsky \etal~\cite{bagon_2011_arxiv}.
We iteratively improve the best solution by
fusing it with proposal solutions.
The inherent difference is how we define the fusion.
We fuse the segmentations according to Eq.~\ref{eq:fusion_move_a}
and optimize the subproblems as described in 
Sec.~\ref{sec:fast_optimization} (see Fig. \ref{fig:move_graph}).

>>>>

Fusion moves as defined in \cite{Lempitsky-2010} work
in the node domain and do not work properly for 
objective functions as Eq.~\ref{eq:nodeproblem} since
the node coloring is ambiguous and has no semantic meaning (see Fig.~\ref{fig:fusion_abiguty}).
In the following, we propose a more suitable fusion move for correlation
clustering which works on the edge domain.
%
Given two proposal solutions $y'$ and $y''$,
$\mathcal{C}$ is the set of edges
which are uncut in $y'$ and $y''$.
%
\begin{align}
\breve{y}    & =  y' \cup y'' \\
\mathcal{C}  & =  \{ (i, j) \quad | \quad \breve{y}_{ij} = 0 \}
\end{align}
%
The fusion move for correlation clustering is solving Eq.~\ref{eq:edgeproblem}
with additional \emph{must-link constraints} for all edges in $\mathcal{C}$.
%
\begin{align}
  y^* &=& \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:fusion_move_a}.\\ 
  s.t & &  y_{ij} = 0 \quad \forall (i, j) \in \mathcal{C} \nonumber
\end{align}
%
By construction, solving Eq.~\ref{eq:fusion_move_a} cannot increase the energy:
\begin{align}
  \sum  w_{ij} \cdot y^*_{ij}  \leq \min \left( \sum  w_{ij} \cdot y'_{ij} ,  \sum   w_{ij} \cdot y''_{ij} \right)
\end{align}
%

Once we know how to fuse segmentations, the actual algorithm 
for correlation clustering 
is simple and similar to the fusion move algorithm of Lempitsky \etal~\cite{bagon_2011_arxiv}.
We iteratively improve the best solution by
fusing it with proposal solutions.
The inherent difference is how we define the fusion.
We fuse the segmentations according to Eq.~\ref{eq:fusion_move_a}.
%
How to optimize  Eq.~\ref{eq:fusion_move_a} in an more efficient way is
described in the following section.

@@

\subsection{Fast Optimization} \label{sec:fast_optimization}

Eq.~\ref{eq:fusion_move_a} can be solved efficiently, since 
we can work on a coarser graph, where all nodes which are connected
via must-link constraints are merged into single nodes.
This coarser graph is constructed by contracting all must-link edges
(see Fig.~\ref{fig:move_graph}).

TODO:
Here we need to proof that it is equivalent 
to optimize the contracted graph !

@@

%-------------------------------------------------------------------------
%\subsection{Fusion Move Solver}
